Consistency scores in text data

### 2020-08-03 outputs/documents/paper.Rmd
recover previous changes in citations in later paragraphs

Given the nature of human languages, some combinations of words tend to occur more frequently than others. Think of 'good', which is more often followed by 'morning', than 'duck'. As such, we could consider English text production as a conditional probability, $\mbox{Pr}(w_{k} | w^{k-1}_{1})$, where $k$ is the number of words in a sentence, $w_{k}$ is the predicted word, and $w^{k-1}_{1}$ is the history of the word occurring in a sequence [@brown1992class]. In this way, the generation of some prediction, $w_{k}$, is based on the history, $w^{k-1}_{1}$. This is the underlying principle of all language models. Essentially, a language model is a probability distribution over sequences of words **(Ke-Li add ref)**. The goal of statistical language modeling is to estimate probability distributions over different linguistic units — words, sentences, and even documents **(Ke-Li add ref)**. However, this is difficult as language is categorical. If we consider each word in a vocabulary as a category, then the dimensionality of language becomes large [@rosenfeld2000two]. The reason that there is scuh a variety of statistical language models is that there are various ways of dealing with this fundamental problem.

@@

Language models underpinned by n-grams are widely applied in text prediction, spelling-correction, and machine translation **(Ke-Li add citation)**. In our application, we use a tri-gram based model to detect both real-word and non-word errors and correct them with the candidate word that has the highest probability. **(TODO: Add more about how we use n-gram once the functions in the app are in place.)**

@@

## Word Embedding Models
Turning words into numeric representations is an important step for machine to understand and process words. The most basic way to turn a word into a vector is one-hot encoding. However, with one-hot encoding, the number of dimensions depends on the size of the vocabulary in a given corpus, and can easily grow to extremely large number of dimensions. Word embedding models work by representing words as numerical dense vectors so the dimensions are reduced by representing the words by features **(Keli add citation)**. Representing words as dense vectors has a long history, but the idea of a dense embedding space for words was first explored in 2003 by @bengio2003neural, and became the prominent driver for word embedding models. In the work of @bengio2003neural, words are encoded into feature vectors that represents different aspects of the word, and each word is associated with a point in a vector space. The words that are similar will have similar representations in a vector space. The similarity of the words, however, does not depend on their spelling or forms, but by their co-occurrence in given contexts. For example, a "cat" is more similar to a "dog" instead of a "car" because we are more likely to talk about cats and dogs in a same conversation. Therefore, the co-occurrence of "cat" and "dog" has higher probability than the co-occurrence of "cat" and a "car". In other words, the distributed representation is learned by the usage of words. This permits words that are used in similar contexts to result in having similar vector representations and their semantic meanings are captured naturally.

@@

Other Word2Vec use cases:
1. Find the most similar words (**Keli add example**)
2. Calculate the similarities between a pair of words (**Keli add example**)
3. Find the odd word in a list of words (**Keli add example**)

@@

In recent years, the use of pre-trained language models has played a significant role in advancing natural language processing (NLP) tasks such as reading comprehension, text generation, and even creative writing **(Ke-Li add ref)**. Pre-trained models remove the need for researchers to prepare training datasets or allot computational resources to the stage of pre-training. Because of this, there has been substantial development in NLP research. OpenAI and Google are two companies that are considered pioneers in artificial intelligence and they both develop pre-trained models that are considered state-of-art; we will discuss the GPT models from OpenAI and the BERT model from Google.

@@

Draw on and cite: 'Language Models are Few-Shot Learners'

OpenAI is an AI research company who openly pursues Artificial General Intelligence—to have machines learn and understand any intellectual tasks as human do **(Ke-Li add ref)**. OpenAI's Generative Pre-Trained Transformer (GPT) models are pre-trained language models that are built based on the transformer architecture. Proposed by @vaswani2017attention from Google in 2017, the Transformer is a novel network architecture for neural network that outperforms RNN-based models in computational efficiency and CNN-based in (something that it doesn't have to look for longer sentences) **(Keli add citation)**. [Write about self-attention layer and how it differentiate from RNN and CNN]. [Paraphrase: The transformer is based solely on attention mechanisms, dispensing with recurrence and convolutions entirely, and has proven extremely effective compared to its predecessor the LSTM].

As of August 2020, OpenAI has three generations of GPT models. GPT was introduced to the public in June, 2018. In the year 2019, its successor GPT-2 was released in three stages. Not long after, in June 2020, OpenAI announced GPT-3, which is the latest GPT model as of July 2020 (@brown2020language). Since the API of GPT-3 at the moment is only available to a few invited users, public applications of GPT-3 are limited. On the other hand, its predecessor, GPT-2, has enabled researchers made breakthroughs in the past years. @alt2019fine extended GPT-2 to relation extraction, which is a task of identifying the relationship between concepts appeared in text. They had shown a milestone in predicting larger range of distinct relation types with higher confidence **(Ke-Li add ref)**. In the field of speech recognition, data scarcity is often an issue due to the difficulty to collect large amount of data from human for learning **(Ke-Li add ref)**. @bird2020lstm employed GPT-2 as part of the solution to generate augmented data for classification. GPT-2 also helped NLP researchers tackling the task of textual paraphrasing **(Ke-Li add ref)**. @hegde2020unsupervised used GPT-2 to build an unsupervised paraphrasing model that is not restricted by domain shift within an article, which is a problem faced by the usual supervised models of paraphrasing.

Most of prior language models are trained in a single domain by supervised learning to perform a single task. OpenAI, on the other hands, train the GPT models to perform many tasks (@radford2019language). GPT-2 is pre-trained on WebText, a comprehensive data set that is scrapped from millions of webpages and contains total of 40 GB of text (@radford2019language). WebText contains all kinds of textual data: dialogues, comments, translations, questions and answers...etc. Such diversity is critical for GPT-2 to be a generalized model that can perform multiple tasks with any domain boundary. However, GPT-2 has the capacity to adapt to a context and generates custom output with the "fine-tuning" technique **(Ke-Li add ref)**. While the pre-training stage is unsupervised and mandatory, fine-tuning is an optional training stage where GPT can learn to perform new tasks in a supervised or reinforced approach (@ziegler2019fine) **(Ke-Li add examples)**. During the pre-training stage, the model is trained to perform text completion task. Give it a prompt to start and it will generate texts that continues what has been started. In the fine-tuning stage, @radford2019language provided additional training data sets for GPT-2 to generate the texts with specific sentiment, or be descriptive, or even summarize the prompt text.

In our project, we aim to use fine-tune GPT-2 to perform a text cleaning task **(Ke-Li add more about how we use fine-tuning to train GPT-2 for text cleaning)**.

A language model is just a probability:
> p(output | input)

To model all NLP tasks, we cab represent them with one extra conditioning:
> p(output | input, task)

OpenAI has made GPT-2 open-sourced through GitHub with Python and TensorFlow. There is also a package in R that wraps the Python code into a R package, created by **(Keli add citation)**. However, the R package only permits exploratory experiments with the pre-trained tasks; fine-tuning for new tasks is not available thorough R.

GPT-3 has over 175 billions of parameters (@brown2020language). The parameters refer to the weights of the connections that enable the neural network to learn **(Keli add citation)**. In the past few years, the growth of parameters of transformer language models expands rapidly. The first generation GPT has 110 million parameters. Released in 2018, GPT-2 has 1.6 billion parameters, and within two years, GPT-3 has grown the size to 175 billion, which is over 100 times than its predecessor. The more parameters there are the better the model can generalize to new tasks; therefore, the increasing amount of parameters accelerates the improvements in text synthesis and downstream NLP tasks **(Keli add citation)**. GPT-3 can learn NLP tasks with just a handful of example; OpenAI refers this ability as few-shot learning, or in-context learning. **(Keli add citation)**. As a result, we only need to provide GPT-3 a few demonstrations for it to learn a new NLP task. For example, OpenAI team trained GPT-3 to generate "news articles". The prompt input is a plausible first sentence of a news story written by human, and the output is an entire news article generated by the model itself. Without the demonstrations, GPT-3 tends to treat the prompts as "tweets" and generate texts that serve as responses or follow-up tweets. The team fed a few previous news articles in the model’s context as demonstrations, and the model learned to produce the outputs that adapt to the style and length of news stories (@brown2020language). This few-shot learning ability indicates that the model can learn like human do and would democratize the use of language models to broader fields.

@@

Because BERT employs the Transformer architecture with the self-attention mechanism, fine-tuning is also available.

>>>>

Given the nature of human languages, some combinations of words tend to occur more frequently than others. Think of 'good', which is more often followed by 'morning', than 'duck'. As such, we could consider English text production as a conditional probability, $\mbox{Pr}(w_{k} | w^{k-1}_{1})$, where $k$ is the number of words in a sentence, $w_{k}$ is the predicted word, and $w^{k-1}_{1}$ is the history of the word occurring in a sequence [@brown1992class]. In this way, the generation of some prediction, $w_{k}$, is based on the history, $w^{k-1}_{1}$. This is the underlying principle of all language models. Essentially, a language model is a probability distribution over sequences of words. The goal of statistical language modeling is to estimate probability distributions over different linguistic units — words, sentences, and even documents (@bengio2003neural). However, this is difficult as language is categorical. If we consider each word in a vocabulary as a category, then the dimensionality of language becomes large [@rosenfeld2000two]. The reason that there is scuh a variety of statistical language models is that there are various ways of dealing with this fundamental problem.

@@

Language models underpinned by n-grams are widely applied in text prediction, spelling-correction, and machine translation  (@brown1992class). In our application, we use a tri-gram based model to detect both real-word and non-word errors and correct them with the candidate word that has the highest probability. **(TODO: Add more about how we use n-gram once the functions in the app are in place.)**

@@

## Word Embedding Models

Turning words into numeric representations is an important step for machine to understand and process words. The most basic way to turn a word into a vector is one-hot encoding. However, with one-hot encoding, the number of dimensions depends on the size of the vocabulary in a given corpus, and can easily grow to extremely large number of dimensions. Word embedding models work by representing words as numerical dense vectors so the dimensions are reduced by representing the words by features (@bengio2003neural). Representing words as dense vectors has a long history, but the idea of a dense embedding space for words was first explored in 2003 by @bengio2003neural, and became the prominent driver for word embedding models. In the work of @bengio2003neural, words are encoded into feature vectors that represents different aspects of the word, and each word is associated with a point in a vector space. The words that are similar will have similar representations in a vector space. The similarity of the words, however, does not depend on their spelling or forms, but by their co-occurrence in given contexts. For example, a "cat" is more similar to a "dog" instead of a "car" because we are more likely to talk about cats and dogs in a same conversation. Therefore, the co-occurrence of "cat" and "dog" has higher probability than the co-occurrence of "cat" and a "car". In other words, the distributed representation is learned by the usage of words. This permits words that are used in similar contexts to result in having similar vector representations and their semantic meanings are captured naturally. As a result, word embedding models are capable of finding the most similar words, calculate the similarities between a pair of words, and find the odd word in a group of words @mikolov2013efficient. The most commonly used word embedding models are Word2Vec and GloVe, which will be described in the following sections.

@@

@@

In recent years, the use of pre-trained language models has played a significant role in advancing natural language processing (NLP) tasks such as reading comprehension, text generation, and even creative writing (@brown2020language). Pre-trained models remove the need for researchers to prepare training datasets or allot computational resources to the stage of pre-training. Because of this, there has been substantial development in NLP research. OpenAI and Google are two companies that are considered pioneers in artificial intelligence and they both develop pre-trained models that are considered state-of-art; we will discuss the GPT models from OpenAI and the BERT model from Google.

@@

OpenAI is an AI research company who openly pursues Artificial General Intelligence — to have machines learn and understand any intellectual tasks as human do **(Cite OpenAI blog)**. OpenAI's Generative Pre-Trained Transformer (GPT) models are pre-trained language models that are built based on the transformer architecture. Proposed by @vaswani2017attention from Google in 2017, the Transformer is a novel network architecture for neural network that outperforms RNN-based and CNN-based models in computational efficiency (@vaswani2017attention). Since the emergence of the Transformer model, most of the representative pre-trained models are built based on this architecture (@hanretty2018comparing). Therefore, we can say that Transformer marks a new era of language models.

As of August 2020, OpenAI has three generations of GPT models. GPT was introduced to the public in June, 2018. In the year 2019, its successor GPT-2 was released in three stages. Not long after, in June 2020, OpenAI announced GPT-3, which is the latest GPT model as of July 2020 (@brown2020language). Since the API of GPT-3 at the moment is only available to a few invited users, public applications of GPT-3 are limited. On the other hand, its predecessor, GPT-2, has enabled researchers made breakthroughs in the past years. @alt2019fine extended GPT-2 to relation extraction, which is a task of identifying the relationship between concepts appeared in text. They had shown a milestone in predicting larger range of distinct relation types with higher confidence (@alt2019fine). In the field of speech recognition, data scarcity is often an issue due to the difficulty to collect large amount of data from human for learning (@bird2020lst). @bird2020lstm employed GPT-2 as part of the solution to generate augmented data for classification. GPT-2 also helped NLP researchers tackling the task of textual paraphrasing. @hegde2020unsupervised used GPT-2 to build an unsupervised paraphrasing model that is not restricted by domain shift within an article, which is a problem faced by the usual supervised models of paraphrasing.

Most of prior language models are trained in a single domain by supervised learning to perform a single task. OpenAI, on the other hands, train the GPT models to perform many tasks (@radford2019language). GPT-2 is pre-trained on WebText, a comprehensive data set that is scrapped from millions of webpages and contains total of 40 GB of text (@radford2019language). WebText contains all kinds of textual data: dialogues, comments, translations, questions and answers...etc. Such diversity is critical for GPT-2 to be a generalized model that can perform multiple tasks with any domain boundary. However, GPT-2 has the capacity to adapt to a context and generates custom output with the "fine-tuning" technique (@ziegler2019fine). While the pre-training stage is unsupervised and mandatory, fine-tuning is an optional training stage where GPT can learn to perform new tasks in a supervised or reinforced approach such as stylistic continuation and summarization (@ziegler2019fine). During the pre-training stage, the model is trained to perform text completion task. Give it a prompt to start and it will generate texts that continues what has been started. In the fine-tuning stage, @radford2019language provided additional training data sets for GPT-2 to generate the texts with specific sentiment, or be descriptive, or even summarize the prompt text.

OpenAI has made GPT-2 open-sourced through Github with Python and Tensorflow. There is also a package in R that wraps the Python code into a R package, created by @keydanaluraschi2019gpt2. However, the R package only permits exploratory experiments with the pre-trained tasks; fine-tuning for new tasks is not available thorough R. In our project, we aim to use fine-tune GPT-2 to perform a text cleaning task **(Ke-Li add more about how we use fine-tuning to train GPT-2 for text cleaning, or try with other task if text cleaning is not achieved)**.

GPT-3 has over 175 billions of parameters (@brown2020language). In the past few years, the growth of parameters of transformer language models expands rapidly. The first generation GPT has 110 million parameters. Released in 2018, GPT-2 has 1.6 billion parameters, and within two years, GPT-3 has grown the size to 175 billion, which is over 100 times than its predecessor. The more parameters there are the better the model can generalize to new tasks; therefore, the increasing amount of parameters accelerates the improvements in text synthesis and downstream NLP tasks (@brown2020language). GPT-3 can learn NLP tasks with just a handful of example; OpenAI refers this ability as few-shot learning, or in-context learning (@brown2020language). As a result, we only need to provide GPT-3 a few demonstrations for it to learn a new NLP task. For example, OpenAI team trained GPT-3 to generate "news articles". The prompt input is a plausible first sentence of a news story written by human, and the output is an entire news article generated by the model itself. Without the demonstrations, GPT-3 tends to treat the prompts as "tweets" and generate texts that serve as responses or follow-up tweets. The team fed a few previous news articles in the model’s context as demonstrations, and the model learned to produce the outputs that adapt to the style and length of news stories (@brown2020language). This few-shot learning ability indicates that the model can learn like human do and would democratize the use of language models to broader fields.

@@

> **(Keli adds example)**

Because BERT employs the Transformer architecture with the self-attention mechanism, fine-tuning is also available.[To be continued...]
